```{r kopts, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
	fig.width = 12,
	fig.height = 8
)
```

```{r libs, include = FALSE, echo = FALSE}
library(knitr)
library(WGCNA)
library(annmatrix)
library(heatmaply)
```

```{r data load, include = FALSE, echo = FALSE}
data_mat = readRDS("../../output/final_annmat.RDS")
#data_mat = data_mat[1:1000,]
```

# Clustering and dendrogram

Sex seems to be the determinng factor in hierarchical clustering.
Neither disease status nor ethnicity seem to be clustered in any meaningful manner.
Also, one sample seems to have a mismatched 'sex' label.

```{r clust, include = TRUE, echo = FALSE}
cor_mat = 1 - abs(cor(data_mat))
cor_dist = dist(cor_mat)
res_hclust = hclust(cor_dist, method = "ward.D")

col_mat = data.frame(
	sex = labels2colors(data_mat$sex),
	disease = labels2colors(data_mat$disease.status),
	ethnicity = labels2colors(data_mat$self.reported.ethnicity))


plotDendroAndColors(res_hclust,
	colors = col_mat,
	dendroLabels = FALSE)
```

# Heatmap

Heatmap with the same clustering. Highly distant groups in rows are separated by sex.

```{r heatmap, include = TRUE, echo = FALSE, warning = FALSE}
cpgsds = rowMeans((data_mat - rowMeans(data_mat))^2)
cpg_hm_no = 200
heatmaply(
	data_mat[order(cpgsds, decreasing = TRUE)[1:cpg_hm_no],],
	labCol = data_mat$sex,
	showticklabels = FALSE,
	Colv = res_hclust,
	k_col = 2, k_row = 2)
```

# PCA

```{r prcomp, include = TRUE, echo = FALSE}
res_prcomp = prcomp(t(data_mat))
```

## Screeplot

```{r screeplot, include = TRUE, echo = FALSE}
no_pcs = 6
barplot((res_prcomp$sdev^2/sum(res_prcomp$sdev^2))[1:no_pcs],
	names.arg= paste0("PC", 1:no_pcs),
	ylab = "Ratio of variance explained")
```

It would take 
`r which(cumsum(res_prcomp$sdev^2) / sum(res_prcomp$sdev^2) > 0.9)[1]`
principal components to capture 90% of variance in the data.

## Pair plot

Sex by colors.
Disease status by shape.
The sample with a mismatched 'sex' label is visible here too.

```{r pairplot, include = TRUE, echo = FALSE}
pairs(res_prcomp$x[,1:5],
	col = labels2colors(data_mat$sex),
	pch = as.numeric(as.factor(data_mat$disease.status)))
```

## PC Heatmap

The heatmap of principal component scroes doesn't reveal much.

```{r pca heatmap, include = TRUE, echo = FALSE, warning = FALSE}
heatmaply(
	t(res_prcomp$x),
	showticklabels = FALSE)
```

## PC relation with metadata {.tabset}

### Sex

```{r pcs and sex, echo = FALSE, include = TRUE}
pairs(res_prcomp$x[,1:5],
	col = labels2colors(data_mat$sex))
```

### Disease Status

```{r pcs and disstat, echo = FALSE, include = TRUE}
pairs(res_prcomp$x[,1:5],
	col = labels2colors(data_mat$disease.status))
```

### Age

The lighter the point, the higher the age.

```{r pcs and age, echo = FALSE, include = TRUE}
color_ramp = colorRampPalette(c("white", "black"))(length(unique(as.numeric(data_mat$age))))
color_ramp = color_ramp[match(as.numeric(data_mat$age), unique(as.numeric(data_mat$age)))]
pairs(res_prcomp$x[,1:5],
	col = color_ramp)
```

### Ethnicity

```{r pcs and etno, echo = FALSE, include = TRUE}
pairs(res_prcomp$x[,1:5],
	col = labels2colors(data_mat$self.reported.ethnicity))
```
