
```{r minfi, echo = FALSE, include = FALSE}
library(minfi)
```

```{r input, echo = FALSE, include = FALSE}
baseDir <- "input/"
targets <- readRDS("output/sample_key.RDS")
targets$Basename <- targets$basename
RGS_file <- "output/RGCS.RDS"
if(!file.exists(RGS_file)){
	met_data <- read.metharray.exp(targets = targets)	
	saveRDS(met_data, RGS_file)
} else {
	met_data <- readRDS(RGS_file)
}
```

# DNA methylation signature of human fetal alcohol spectrum disorder

- GEO ID: GSE80261

The data describes genome-wide DNA methylation patterns of
buccal epithelial cells taken from 206 children, 96 of which are controls
and 110 are cases. DNA was analysed using the Illumina450k array.

## The data

Dimensions of the data:

```{r dims, echo = TRUE, include = TRUE)
dim(met_data)
```

That is, there are `r ncol(met_data)` samples and 
`r nrow(met_data)` probes.


## Info on some samples

```{r summary, echo = FALSE, include = FALSE}
head(pData(met_data))
```

## Info on some probes

```{r summary, echo = FALSE, include = FALSE}
head(getProbeInfo(met_data))
```
