```{r libs, echo = FALSE, include = FALSE}
library(annmatrix)
library(ggplot2)
library(reshape2)
library(basetheme)
library(MASS)
library(VennDiagram)
```


```{r data load, echo = FALSE, include = FALSE}
met_annmat = readRDS("../output/step1_annmat.RDS")
```

# Outlier removal {.tabset}

## Mean distance from other samples {.tabset}

### Initial

```{r mean dists, echo = FALSE, include = TRUE}
sd_cutoff = 1.5

dist_outliers <- c()

sample_dists <- as.matrix(dist(t(met_annmat)))
mean_dists <- scale(colMeans(sample_dists))

hist(mean_dists, breaks=15, col = "green")
abline(v = sd_cutoff, col = "red")

dist_outliers <- c(dist_outliers, colnames(met_annmat)[mean_dists > sd_cutoff])

met_annmat <- met_annmat[, mean_dists < sd_cutoff]
```

### After 1st round

After removal of outliers

```{r mean dists after removal, echo = FALSE, include = TRUE}

sample_dists <- as.matrix(dist(t(met_annmat)))
mean_dists <- scale(colMeans(sample_dists))

hist(mean_dists, breaks=15, col = "green")
#abline(v = sd_cutoff, col = "red")

```

Number of samples left: `r ncol(met_annmat)`


## IAC {.tabset}

### Correlation matrix

```{r correlation matrix, echo = FALSE, include= TRUE}

IAC_outliers <- c()

met_annmat = readRDS("../output/step1_annmat.RDS")

cor_mat <- cor(met_annmat)
ggplot(melt(cor_mat)) +
	geom_tile(aes(x = Var1, y = Var2, fill = value)) +
	theme(axis.text.x=element_blank(), axis.text.y=element_blank()) +
	ylab("") + xlab("")
```

### Mean correlations {.tabset}

Row means of the correlation matrix

#### Initial

```{r mean cor init, echo = FALSE, include = TRUE}
cor_means <- colMeans(cor_mat)

basetheme(basetheme_dark())
hist(cor_means, col = "red")
```

#### Iteration No. 1

```{r mean cor it1, echo = FALSE, include = TRUE}
sd_cutoff = -2

cor_mat <- cor(met_annmat)
cor_means <- scale(colMeans(cor_mat))
stopifnot(sum(rownames(cor_means) != colnames(met_annmat)) == 0)
met_annmat <- met_annmat[,cor_means > sd_cutoff]

IAC_outliers <- c(IAC_outliers, colnames(met_annmat)[cor_means < sd_cutoff])

hist(cor_means, col = "red", main = "standard units of IACs", xlab = "")
abline(v = sd_cutoff, col = "blue")

```

#### Iteration No. 2

```{r mean cor it2, echo = FALSE, include = TRUE}
sd_cutoff = -2

cor_mat <- cor(met_annmat)
cor_means <- scale(colMeans(cor_mat))
stopifnot(sum(rownames(cor_means) != colnames(met_annmat)) == 0)
met_annmat <- met_annmat[,cor_means > sd_cutoff]

IAC_outliers <- c(IAC_outliers, colnames(met_annmat)[cor_means < sd_cutoff])

hist(cor_means, col = "red", main = "standard units of IACs", xlab = "")
abline(v = sd_cutoff, col = "blue")

```

#### Final

```{r mean cor final, echo = FALSE, include = TRUE}
sd_cutoff = -2

cor_mat <- cor(met_annmat)
cor_means <- scale(colMeans(cor_mat))
stopifnot(sum(rownames(cor_means) != colnames(met_annmat)) == 0)
met_annmat <- met_annmat[,cor_means > sd_cutoff]

IAC_outliers <- c(IAC_outliers, colnames(met_annmat)[cor_means < sd_cutoff])

hist(cor_means, col = "red", main = "standard units of IACs", xlab = "")

```


Number of samples left: `r ncol(met_annmat)`

## Intersection of outliers between methods

```{r, intersect, echo = FALSE, include = TRUE}
suppressMessages(suppressWarnings(
draw.pairwise.venn(area1 = length(IAC_outliers),
	area2 = length(dist_outliers),
	cross.area = length(intersect(dist_outliers,
		IAC_outliers)),
	category = c("IAC", "Distances"),
	fill = c("light blue", "pink"),
	lty = "blank")
))
```

Both methods capture completely different outliers.

# Quality control

## Distribution of methylation in different regions

As expected, CpG seas and shelves have high methylation values while
islands and shores have low methylation.

```{r region densities, echo = FALSE, include = TRUE}
cpg_means <- rowMeans(met_annmat)
names(cpg_means) <- met_annmat@Relation_to_Island
cpg_shore <- cpg_means[names(cpg_means) %in% c("N_Shore", "S_Shore")]
cpg_shelf <- cpg_means[names(cpg_means) %in% c("N_Shelf", "S_Shelf")]
cpg_island <- cpg_means[names(cpg_means) == "Island"]
cpg_sea <- cpg_means[names(cpg_means) == "OpenSea"]

basetheme(basetheme_brutal())

plot(density(cpg_sea), col = "blue", main = NA, cex= 2)
lines(density(cpg_island), col = "palegoldenrod", cex = 2)
lines(density(cpg_shore), col = "palegreen", cex = 2)
lines(density(cpg_shelf), col = "purple4", cex = 2)

legend(x = 0.3, y = 3,
	legend = c("sea", "island", "shore", "shelf"),
	col = c("blue", "palegoldenrod", "palegreen", "purple4"),
	lty = 1
)
```

## Multidimensional scaling {.tabset}

Samples divide based on sex nicely when using both MDS methods.

### PCA

```{r pca, echo = FALSE, include = TRUE}
met_pca <- prcomp(t(met_annmat))$x
plot(met_pca[,1:2], col = as.factor(met_annmat$sex), main = "PCA, coloured by sex")
```

### Sammon map

```{r sammon, echo = FALSE, include = TRUE}
met_sample_dist = dist(t(met_annmat))
met_sammon <- sammon(met_sample_dist)
plot(met_sammon$points, col = as.factor(met_annmat$sex), main = "Sammon map, coloured by sex")
```

