```{r libsnopts, echo = FALSE, include = FALSE}
library(annmatrix)
library(knitr)
```

```{r data load, echo = FALSE, include = FALSE}
data_mat = readRDS("../../output/final_annmat.RDS")
data_mat = data_mat[1:1000,]
gc()
```

```{r ttest, echo = FALSE, include = FALSE}
tests = apply(data_mat, 1,
	function(x) t.test(x[data_mat$disease.status == "Control"],
		x[data_mat$disease.status != "Control"]))
t.results = data.frame(pvalues = sapply(tests, function(x) x$p.value),
	effect = sapply(tests, function(x) diff(x$estimate)))
sig_names = rownames(t.results[order(t.results$pvalues),])[1:5]
```

# Comparing FASD and healthy CpG spots {.tabset}

5 most significant positions. Colored by disease status

```{r plot5, echo = FALSE, include = TRUE, results = "asis"}
sig_vals = data_mat[match(sig_names, data_mat@Name),
	order(data_mat$disease.status)]
cat("\n")
for(i in 1:5){
	cat(paste0("## ",rownames(sig_vals)[i], "\n"))
	cat("\n")
	plot(sig_vals[i,], ylab = "methylation", pch = 21,
		col = as.factor(sig_vals$disease.status))
	cat("\n")
	cat("\n")
}
```

# P value table

```{r ptab, echo = FALSE, include = TRUE}
p_table = data.frame(
	"Significance" = c("α = 0.1", "α = 0.05",
		"α = 0.01", "FDR = 0.05", "Bonf = 0.05"),
	"No. of genes" = c(
		sum(t.results$pvalues < 0.1),
		sum(t.results$pvalues < 0.05),
		sum(t.results$pvalues < 0.01),
		sum(p.adjust(t.results$pvalues,
			method = "fdr") < 0.05),
		sum(p.adjust(t.results$pvalues,
			method = "bonferroni") < 0.05)))
kable(p_table)
```

# P value histogram

Seems liek there aren't many discoveries. Using a significance level
cutoff directly might produce many false discoveries, so a further p
value adjustment is needed.

```{r phist, echo = FALSE, include = TRUE}
hist(t.results$pvalues, main = "P values", breaks = 80)
```

# Volcano plot

Red and yellow lines mark significance levels 0.05 and 0.01.

```{r vvolcplot, echo = FALSE, include = TRUE}
plot(t.results$effect, -log(t.results$pvalues),
	xlab = "effect size", ylab = "-log P value")
abline(h = -log(0.05), col = "red")
abline(h = -log(0.01), col = "yellow")
```

# Manhattan plot

```{r manhattan, echo = FALSE, include = TRUE}
plot(y = -log(t.results$pvalues[data_mat@chr == "chr1"]),
	x = data_mat@pos[data_mat@chr == "chr1"])
abline(h = -log(0.05), col = "red")
abline(h = -log(0.01), col = "orange")
```

# Linear model

```{r anova, echo = FALSE, include = TRUE}
anova.results = lapply(1:nrow(data_mat), function(x){
	#lm0 = lm(data_mat[x,] ~ as.factor(data_mat$sex))
	#lm1 = lm(data_mat[x,] ~ as.factor(data_mat$sex) + as.numeric(data_mat$age))
	lm0 = lm(data_mat[x,] ~ as.factor(data_mat$sex))
	lm1 = lm(data_mat[x,] ~ as.factor(data_mat$sex) + as.numeric(data_mat$age))
	return(c(anova(lm0, lm1)$"Pr(>F)"[2],
		lm1$coefficients[3]))
})

t.results = Reduce(rbind, anova.results)
colnames(t.results) = c("pvalues", "effect")
rownames(t.results) = rownames(data_mat)
t.results = as.data.frame(t.results)

```

# Comparing FASD and healthy CpG spots {.tabset}

5 most significant positions. Colored by disease status

```{r lin plot5, echo = FALSE, include = TRUE, results = "asis"}
sig_vals = data_mat[match(sig_names, data_mat@Name),
	order(data_mat$disease.status)]
cat("\n")
for(i in 1:5){
	cat(paste0("## ",rownames(sig_vals)[i], "\n"))
	cat("\n")
	plot(sig_vals[i,], ylab = "methylation", pch = 21,
		col = as.factor(sig_vals$disease.status))
	cat("\n")
	cat("\n")
}
```

# P value table

```{r lin ptab, echo = FALSE, include = TRUE}
p_table = data.frame(
	"Significance" = c("α = 0.1", "α = 0.05",
		"α = 0.01", "FDR = 0.05", "Bonf = 0.05"),
	"No. of genes" = c(
		sum(t.results$pvalues < 0.1),
		sum(t.results$pvalues < 0.05),
		sum(t.results$pvalues < 0.01),
		sum(p.adjust(t.results$pvalues,
			method = "fdr") < 0.05),
		sum(p.adjust(t.results$pvalues,
			method = "bonferroni") < 0.05)))
kable(p_table)
```

# P value histogram

Seems liek there aren't many discoveries. Using a significance level
cutoff directly might produce many false discoveries, so a further p
value adjustment is needed.

```{r lin phist, echo = FALSE, include = TRUE}
hist(t.results$pvalues, main = "P values", breaks = 80)
```

# Volcano plot

Red and yellow lines mark significance levels 0.05 and 0.01.

```{r lin vvolcplot, echo = FALSE, include = TRUE}
plot(t.results$effect, -log(t.results$pvalues),
	xlab = "effect size", ylab = "-log P value")
abline(h = -log(0.05), col = "red")
abline(h = -log(0.01), col = "yellow")
```

# Manhattan plot

```{r lin manhattan, echo = FALSE, include = TRUE}
plot(y = -log(t.results$pvalues[data_mat@chr == "chr1"]),
	x = data_mat@pos[data_mat@chr == "chr1"])
abline(h = -log(0.05), col = "red")
abline(h = -log(0.01), col = "orange")
```
